---
title: "Tidy and filter data: Food Balance Sheets"
author: "Gordon Blasco, Danielle Ferraro"
date: "9/18/2019"
output: html_document
---

## Overview

This document contains code to tidy data for Track 2 of the Clean Seafood project. We use FAO Food Balance Sheets (FBS) for food supply data. Two products are created here: a complete FBS dataset that has been tidied, and one that has been both tidied and filtered. The filtering steps taken were:

- Remove all aggregated food items (e.g. Animal products) leaving only individual food items
- Remove all aggregated regions (e.g. continents) leaving only individual countries
- Remove Yugoslavia Socialist Federal Republic countries due to discontinuities between the former and current time series:
  - Slovenia
  - Croatia
  - Bosnia and Herzegovina
  - Macedonia
  - Serbia and Montenegro (union and separate)
- Remove former Soviet Union countries due to discontinuities between the former and current time series:
  - Armenia
  - Azerbaijan
  - Belarus
  - Estonia
  - Georgia
  - Kazakhstan
  - Kyrgyzstan
  - Latvia
  - Lithuania
  - Moldova
  - Russia
  - Tajikistan
  - Turkmenistan
  - Ukraine
  - Uzbekistan
- Remove Luxembourg (incomplete time series; data begins 2000) 
- Remove former Sudan (only two years of data)

## Setup
```{r}
library(here)
library(tidyverse)
library(janitor)
library(readxl)

source(here("src", "directories.R")) # Sets file directories 
```


## Import data
```{r}
# Food balance sheets
food_balance <- read_csv(file.path(dir_raw_data, "FAO/food_balance/FoodBalanceSheets_E_All_Data_NOFLAG.csv"))

area_code <- food_balance %>% 
  distinct(`Area`, `Area Code`) %>% 
  clean_names()

item_code <- food_balance %>% 
  distinct(`Item`, `Item Code`) %>% 
  clean_names()

element_code <- food_balance %>% 
  distinct(`Element`, `Element Code`) %>% 
  clean_names()

code_item_group <- read_csv(file.path(dir_raw_data, "FAO", "food_balance", "FoodBalanceSheets_Metadata_ItemGroupCode.csv"), locale = locale(encoding = "latin1")) %>% 
  clean_names()
```

## Tidy data
```{r}
# Food balance sheets
clean_food_balance <- food_balance %>% 
  gather(
    key = "year",
    value = "quantity",
    c(8:60)
  ) %>% 
  clean_names() %>% 
  mutate(year = parse_number(year))
```

## Identify data to filter out
```{r}
# Define list of USSR and successor nations
former_ussr <-   
  c("Armenia",
  "Azerbaijan",
  "Belarus",
  "Estonia",
  "Georgia",
  "Kazakhstan",
  "Kyrgyzstan",
  "Latvia",
  "Lithuania",
  "Republic of Moldova",
  "Russian Federation",
  "Tajikistan",
  "Turkmenistan",
  "Ukraine",
  "Uzbekistan")

length(former_ussr)
area_code %>% filter(area %in% former_ussr) %>% nrow()

# Define list of Yugoslavia SFR and successor nations
former_yugoslavia <- 
  c("Slovenia",
  "Croatia",
  "Bosnia and Herzegovina",
  "The former Yugoslav Republic of Macedonia",
  "Serbia and Montenegro",
  "Serbia",
  "Montenegro")

length(former_yugoslavia)
area_code %>% filter(area %in% former_yugoslavia) %>% nrow()

# Combine countries to remove in one dataframe
countries_to_remove <- tibble(area = c(former_ussr, 
                                       former_yugoslavia, 
                                       "Luxembourg",
                                       "Sudan (former)")) %>% 
  left_join(area_code, by = "area")
filter(countries_to_remove, is.na(area_code)) # Check for NAs
```

## Filter data
```{r}
filtered_food_balance <- clean_food_balance %>% 
  filter(!area_code %in% countries_to_remove$area_code, # Remove countries as defined above
         area_code < 5000, # Remove aggregated geographic regions
         !item_code %in% unique(code_item_group$item_group_code)) # Remove item groups (keep individual items only)
```

## Write the csvs
```{r}
# Unfiltered food balance sheets and metadata
write_csv(clean_food_balance, file.path(dir_data_bottlenecks, "FAO", "food_balances", "food_balance_tidy.csv"))
write_csv(area_code, file.path(dir_data, "bottlenecks_data", "FAO", "food_balances", "area_code.csv"))
write_csv(item_code, file.path(dir_data, "bottlenecks_data", "FAO", "food_balances", "item_code.csv"))
write_csv(element_code, file.path(dir_data, "bottlenecks_data", "FAO", "food_balances", "element_code.csv"))

# Filtered food balance sheets
write_csv(filtered_food_balance, file.path(dir_data_bottlenecks, "FAO", "food_balances", "food_balance_tidy_filtered.csv"))
```

