---
title: "Identify proximate drivers of food supply declines"
author: "Danielle Ferraro"
date: "`r Sys.Date()`"
output: pdf_document
---

# Overview

1) Perform Spearman's rank correlations on supply and the components of supply (production, imports, exports, etc) to identify which component was the proximate driver (i.e., dominant correlate) of supply during declines. 

2) Identify if, during production-associated supply declines, there were also declines in absolute production, harvestable area, or yield.

# Setup

Load packages.
```{r}
library(here)
library(tidyverse)
library(janitor)
library(vroom)
library(ggpubr)
library(paletteer)
library(patchwork)
library(broom)
```

Set directories.
```{r}
source(here("src", "directories.R"))
```

# Load data 

```{r}
# Tidy food balance data
food_balance <- vroom(file.path(dir_data_bottlenecks, "FAO", "food_balances", "food_balance_tidy_filtered.csv"), delim = ",", locale = locale(encoding = "latin1"))

# Raw harvestable area/yield, from FAO crops dataset
crops <- vroom(file.path(dir_raw_data, "FAO", "crops_livestock", "Production_Crops_Livestock_E_All_Data_(Normalized).csv"), delim = ",", locale = locale(encoding = "latin1")) %>% 
  clean_names()

# Key to map FBS items to crops items
fbs_to_crops_key <- vroom(file.path(dir_raw_data, "FAO", "crops_livestock", "MapSPAM_to_FAO_v2.csv"), delim = ",", locale = locale(encoding = "latin1"))

# Food supply declines list
declines <- read_csv(file.path(dir_data_bottlenecks, "declines", "characterized_declines_2021-08-13.csv"))

# Coincident events list
coincident_events <- read_csv(file.path(dir_data_bottlenecks, "declines", "declines_w_coincident_inclines.csv"))

# Time series with food supply declines
ts_w_declines <- read_csv(file.path(dir_data_bottlenecks, "declines", "ts_w_declines_2021-08-13.csv"))
```

# Prep data

Subset necessary data:  
* 511 - population
* 5511- production
* 5611 - import quantity
* 5911 - export quantity
* 645 - food supply
* 5301 - total supply
* 5521 - feed  
* 5527 - seed
* 5072 - stock variation
and harmonize units.
```{r}
distinct(food_balance, element, element_code, unit)

food_balance_per_cap <- food_balance %>% 
  
  # Select necessary components of supply
  filter(element_code %in% c(511, 5511, 5611, 5911, 645, 5301, 5521, 5527, 5072)) %>% 
  mutate(quantity = case_when(unit == "1000 tonnes" ~ quantity*1000*1000, # Convert all quantities to kg
                              unit == "1000 persons" ~ quantity*1000, # Multiply population by the unit scaling factor
                              TRUE ~ quantity)) %>% 
  
  # Convert production, imports, exports, and total supply to per capita quantities (food supply is already per capita)
  group_by(area_code, area, year) %>% 
  mutate(quantity = if_else(item == "Population" | element_code == 645, 
                            quantity,
                            quantity/quantity[item == "Population"])) %>% 
  ungroup() %>% 
  
  # Convert to wide format for correlation
  filter(item != "Population") %>% 
  select(-element_code, -unit) %>% 
  pivot_wider(names_from = element, values_from = quantity) %>% 
  
  # Clean up column names
  clean_names() %>% 
  rename(prod_per_cap = production,
         import_per_cap = import_quantity,
         export_per_cap = export_quantity,
         food_supply_per_cap = food_supply_quantity_kg_capita_yr,
         feed_per_cap = feed,
         seed_per_cap = seed,
         stock_var_per_cap = stock_variation,
         total_supply_per_cap = domestic_supply_quantity) %>% 
  
  # Filter to only the years within decline events
  right_join(
    ts_w_declines %>% 
      filter(!is.na(decline_id)) %>%
      rename(food_supply_per_cap = quantity)
    ) %>% # nrow = 30509

  select(decline_id, year, ends_with("per_cap")) %>% 
  arrange(decline_id, year) 
```

# Explore missing data

Though supply is a derived variable, not all supply declines have associated production/imports/exports/etc time series. Some data is completely missing, while other time series are made up of all zeroes. Taking a peek at how much data is missing...

```{r}
food_balance_per_cap %>% 
  group_by(decline_id) %>% 
  naniar::miss_var_summary()

missing_summary <- ts_w_declines %>% 
  filter(!is.na(decline_id)) %>%
  rename(supply_per_cap = quantity) %>%
  left_join(food_balance_per_cap) %>% 
  select(decline_id, supply_per_cap, prod_per_cap, export_per_cap, import_per_cap, seed_per_cap, feed_per_cap) %>%
  group_by(decline_id) %>% 
  naniar::miss_var_summary() %>% 
  mutate(na = if_else(pct_miss == 100, TRUE, FALSE)) %>% 
  ungroup()

missing_summary %>% 
  filter(na == TRUE) %>% 
  count(variable) %>% 
  mutate(prop = n/nrow(declines)) %>% 
  arrange(desc(prop))
```

# Plot supply/supply element time series

Warning: takes a few minutes to run.
```{r, eval = FALSE}
supply_component_ts_plots <- ts_w_declines %>% 
  filter(!is.na(decline_id)) %>%
  rename(supply_per_cap = quantity) %>%
  left_join(food_balance_per_cap) %>% 
  pivot_longer(c("supply_per_cap",
                 "prod_per_cap",
                 "export_per_cap",
                 "import_per_cap",
                 "seed_per_cap",
                 "feed_per_cap"), names_to = "supply_element", values_to = "quantity") %>% 
  group_split(decline_id) %>% 
  map(
    ~ggplot(., aes(x = year, y = quantity, color = supply_element, group = supply_element)) +
      geom_point() +
      geom_line() +
      labs(x = NULL,
           y = "Quantity (kg/capita/yr)",
           color = "Supply element",
           title = paste0(.$area, ": Decline in ", .$item),
           subtitle = paste0("Decline ID#: ", .$decline_id)) +
      cowplot::theme_half_open()
  )

pdf(here("output", "figs", "supply_component_time_series.pdf"), height = 11, width = 8.5)
ggarrange(plotlist = supply_component_ts_plots, ncol = 1, nrow = 3, align = "v")
dev.off()
```

# Calculate ratio of food supply to total supply

Should be close to 1

```{r}
(food_supply_vs_total_supply <- food_balance_per_cap %>% 
   mutate(ratio = food_supply_per_cap/total_supply_per_cap) %>% # NAs, NaNs, Infs...
   group_by(decline_id) %>% 
   summarize(across(where(is.numeric), mean))
)

summary(food_supply_vs_total_supply$ratio)
plot(density(na.omit(food_supply_vs_total_supply$ratio)))
plot(food_supply_vs_total_supply$food_supply_per_cap, food_supply_vs_total_supply$total_supply_per_cap, cex = 0.3)
abline(a = 0, b = 1)

summary(na.omit(food_supply_vs_total_supply$ratio))
```

# Correlations

```{r}
# Append per cap trade data to supply time series (for years within a decline event only) and convert to list
food_balance_per_cap_list <- food_balance_per_cap %>% 
  group_split(decline_id, .keep = FALSE) %>% # The list element index lines up with the decline_id here, but I should write something to ensure this
  map(~column_to_rownames(., "year"))

# Spearman's function
compute_trade_correlations <- function(x_df){
  cor.matrix <- cor(x_df, method = "spearman", use = "pairwise.complete.obs")
  cor.matrix["food_supply_per_cap", 
             !(colnames(cor.matrix) == "food_supply_per_cap")]
  }

# Map correlation function over each country-item ts and put results into data frame
cor_results <- map(food_balance_per_cap_list, compute_trade_correlations) %>% 
  bind_rows(.id = "decline_id") %>%  # Map list element number to a new column
  mutate(decline_id = as.integer(decline_id))

# Identify dominant correlate(s) per decline, i.e., the supply component(s) with abs(rho) >= 0.9, else the component(s) with max(rho).
# We decided to remove seed, feed, and stock variation from the analysis to focus only on production and trade

rho_thresh <- 0.9

main_correlates <- cor_results %>% 
  filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Unrecovered declines only
  # drop_na(ends_with("per_cap")) %>% # Remove declines that have no rho value for ANY correlate (this is due to missing data) -- decided against this, very few have complete data
  select(-total_supply_per_cap,
         -feed_per_cap,
         -seed_per_cap, # Seed positively correlated with food supply
         -stock_var_per_cap) %>% # Stock variation very weakly correlated to supply 
  pivot_longer(cols = ends_with("per_cap"), names_to = "element", values_to = "rho") %>% 
  mutate(rho = abs(rho)) %>% # Take absolute value of rho
  group_by(decline_id) %>% 
  summarize(dominant_correlate = if_else(max(na.omit(rho)) >= rho_thresh, 
                                         paste(element[which(rho >= rho_thresh)], collapse = ", "),
                                         paste(element[which(rho == max(na.omit(rho)))], collapse = ", "))) %>%
  mutate(dominant_correlate = if_else(dominant_correlate == "", NA_character_, dominant_correlate)) %>% 
  ungroup() %>% 
  left_join(select(declines, decline_id, area, area_code, item, item_code, food_group, sector), by = "decline_id") # Rejoin food data

# How many declines had no dominant correlate? (because of missing data or 0 data)
main_correlates %>%
  filter(is.na(dominant_correlate)) # n = 25

# Explore results

filter(main_correlates, is.na(dominant_correlate))
sort(table(main_correlates$dominant_correlate), decreasing = TRUE)
sort(prop.table(table(main_correlates$dominant_correlate)), decreasing = TRUE)
filter(main_correlates, str_detect(dominant_correlate, "feed_per_cap")) %>% arrange(sector) %>% View()

# Is seed positively correlated most of the time?
hist(cor_results$seed_per_cap) # Yes - will remove from analysis
hist(cor_results$feed_per_cap) # Feed too
```

# Extract numbers 

Pull numbers for use in manuscript.
```{r}
abs(cor_results) %>% 
  filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
  summary()

abs(cor_results) %>% 
  filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
  pull(prod_per_cap) %>% 
  density(na.rm = TRUE)

main_correlates %>% 
  drop_na(dominant_correlate) %>% # Remove the 25 declines with missing data first
  filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
  separate_rows(dominant_correlate, sep = ", ") %>% 
  count(dominant_correlate) %>% 
  mutate(prop = n/sum(n)) %>% 
  arrange(desc(prop))

main_correlates %>% 
  drop_na(dominant_correlate) %>% # Remove the 25 declines with missing data first
  filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
  separate_rows(dominant_correlate, sep = ", ") %>% 
  count(food_group, dominant_correlate) %>% 
  group_by(food_group) %>% 
  mutate(prop = n/sum(n)) %>% 
  ungroup() %>% 
  arrange(food_group, desc(prop)) %>% 
  print(n = Inf)
```


# Plot correlation results

## Density of Spearman's rho values across dominant correlates
```{r}
# pal <- "beyonce::X6"
pal <- "beyonce::X1"

(a <- abs(cor_results) %>% 
    filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
    select(-total_supply_per_cap, -stock_var_per_cap, -seed_per_cap, -feed_per_cap) %>% 
    pivot_longer(cols = where(is.double), names_to = "element", values_to = "rho") %>% 
    
    ggplot(aes(x = rho, fill = factor(element, levels = c("prod_per_cap", "import_per_cap", "export_per_cap")), color = factor(element, levels = c("prod_per_cap", "import_per_cap", "export_per_cap")))) +
    geom_density(size = 1.25, alpha = 0.1) +
    geom_vline(xintercept = 0.9, linetype = "dashed") +
    scale_color_manual(values = c(paletteer_d(pal)[2:4]), 
                       labels = c("export_per_cap" = "Exports", 
                                  "import_per_cap" = "Imports",
                                  "prod_per_cap" = "Production")) +
    scale_fill_manual(values = c(paletteer_d(pal)[2:4]), 
                      labels = c("export_per_cap" = "Exports", 
                                 "import_per_cap" = "Imports",
                                 "prod_per_cap" = "Production")) +
    scale_x_continuous(expand = expansion(mult = c(0, 0))) +
    scale_y_continuous(expand = expansion(mult = c(0, 0))) +
    labs(x = "Absolute value of Spearman's rho",
         y = "Density",
         color = bquote(Supply~element~(kg~capita^-1~yr^-1)),
         fill = bquote(Supply~element~(kg~capita^-1~yr^-1))) +
    cowplot::theme_half_open() +
    theme(legend.position = c(0.03, 0.85))
)
```

## Barplot of counts per dominant correlate
```{r}
(b <- main_correlates %>% 
    filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
    drop_na(dominant_correlate) %>% 
    separate_rows(dominant_correlate, sep = ", ") %>% 

    ggplot(aes(x = fct_infreq(dominant_correlate), fill = fct_infreq(dominant_correlate))) +
    geom_bar(aes(y = (..count..)/sum(..count..))) + # Plots proportions

    labs(y = "Percentage of supply declines",
         x = bquote(Dominant~correlate~(kg~capita^-1~yr^-1)),
         fill = "Dominant correlate") +
    scale_fill_manual(values = c(paletteer_d(pal)[2:5]),
                      labels = c("export_per_cap" = "Exports",
                                 "import_per_cap" = "Imports",
                                 "prod_per_cap" = "Production")
    ) +
    scale_x_discrete(labels = c("export_per_cap" = "Exports",
                                "import_per_cap" = "Imports",
                                "prod_per_cap" = "Production"),
                     expand = expansion(mult = c(0.17, 0.17))) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       expand = expansion(mult = c(0, 0))) +
    cowplot::theme_half_open() +
    theme(legend.position = c(0.55, 0.85), 
          plot.margin = unit(c(7,17,7,7), "pt"))
)
```

## Proportional barplot of dominant correlates per food group
```{r}
# Define vector for ordering food groups in the legend
food_group_order <- c("Starchy roots", "Cereals and pulses", "Fruits and vegetables", 
                      "Vegetable oils and oilcrops", "Sugar and sugar crops", "Alcoholic beverages",
                      "Meat and offals", "Milk, eggs, and animal fats", "Seafood", 
                      "Other")

(c <- main_correlates %>% 
    filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
    drop_na(dominant_correlate) %>% 
    separate_rows(dominant_correlate, sep = ", ") %>% 

    ggplot(aes(y = factor(food_group, levels = rev(food_group_order)), fill = fct_infreq(dominant_correlate))) +
    geom_bar(position = position_fill(reverse = TRUE)) +
    labs(x = "Percentage of supply declines",
         y = NULL,
         fill = bquote(Dominant~correlate~(kg~capita^-1~yr^-1))) +
    scale_fill_manual(values = paletteer_d(pal)[2:5],
                           labels = c("export_per_cap" = "Exports",
                                      "import_per_cap" = "Imports",
                                      "prod_per_cap" = "Production",
                                      "feed_per_cap" = "Feed",
                                      "seed_per_cap" = "Seed")) +
    scale_x_continuous(label = scales::percent_format(accuracy = 1),
                       expand = expansion(mult = c(0, 0))) +
    cowplot::theme_half_open() +
    theme(plot.margin = unit(c(7,17,7,7), "pt"))
)

# Version with icons!
library(png) # For importing pngs
library(grid) # For adding pngs to ggplot
# Load icons, and set X position for plotting along the Y axis.
apple <- rasterGrob(readPNG(here("output", "figs", "icons", "apple_black.png")))
sugarcane <- rasterGrob(readPNG(here("output", "figs", "icons", "sugarcane_black.png")))
beer <- rasterGrob(readPNG(here("output", "figs", "icons", "beer_black.png")))
milk <- rasterGrob(readPNG(here("output", "figs", "icons", "milk_black.png")))
olivebranch <- rasterGrob(readPNG(here("output", "figs", "icons", "olivebranch_black.png")))
grouper <- rasterGrob(readPNG(here("output", "figs", "icons", "grouper.png")))
rice <- rasterGrob(readPNG(here("output", "figs", "icons", "rice.png")))
cassava <- rasterGrob(readPNG(here("output", "figs", "icons", "cassava.png")))
cow <- rasterGrob(readPNG(here("output", "figs", "icons", "cow.png")))

icon_xmin <- -0.045
icon_xmax <- -0.015

(c_with_icons <- main_correlates %>% 
    filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
    drop_na(dominant_correlate) %>% 
    separate_rows(dominant_correlate, sep = ", ") %>% 

    ggplot(aes(y = factor(food_group, levels = rev(food_group_order)), fill = fct_infreq(dominant_correlate))) +
    geom_bar(position = position_fill(reverse = TRUE)) +
    annotation_custom(grob = cassava, xmin = icon_xmin, xmax = icon_xmax, ymin = 9.6, ymax = 10.4) +
    annotation_custom(grob = rice, xmin = icon_xmin, xmax = icon_xmax, ymin = 8.6, ymax = 9.4) +
    annotation_custom(grob = apple, xmin = icon_xmin, xmax = icon_xmax, ymin = 7.75, ymax = 8.25) +
    annotation_custom(grob = olivebranch, xmin = icon_xmin, xmax = icon_xmax, ymin = 6.5, ymax = 7.5) +
    annotation_custom(grob = sugarcane, xmin = icon_xmin, xmax = icon_xmax, ymin = 5.5, ymax = 6.5) +
    annotation_custom(grob = beer, xmin = icon_xmin, xmax = icon_xmax, ymin = 4.25, ymax = 5.75) +
    annotation_custom(grob = cow, xmin = icon_xmin, xmax = icon_xmax, ymin = 3.75, ymax = 4.25) +
    annotation_custom(grob = milk, xmin = icon_xmin, xmax = icon_xmax, ymin = 2.25, ymax = 3.75) +
    annotation_custom(grob = grouper, xmin = icon_xmin, xmax = icon_xmax, ymin = 1.25, ymax = 2.75) +
    coord_cartesian(clip = "off") +
    labs(x = "Percentage of supply declines",
         y = NULL,
         fill = bquote(Dominant~correlate~(kg~capita^-1~yr^-1))) +
    scale_fill_manual(values = paletteer_d(pal)[2:5],
                           labels = c("export_per_cap" = "Exports",
                                      "import_per_cap" = "Imports",
                                      "prod_per_cap" = "Production",
                                      "feed_per_cap" = "Feed",
                                      "seed_per_cap" = "Seed")) +
    scale_x_continuous(label = scales::percent_format(accuracy = 1),
                       expand = expansion(mult = c(0, 0))) +
    cowplot::theme_half_open() +
    theme(plot.margin = unit(c(7,17,7,7), "pt"),
          legend.position = "top",
          axis.text.y = element_text(margin = margin(r = 25))) # Adding room for icons
)

# Put everything together:
library(patchwork)
a / (b + c + theme(legend.position = "none")) & plot_annotation(tag_levels = "a")
ggsave(here("output", "figs", "declines_ms", "fig_correlation_results.png"), height = 6, width = 8, dpi = 600, scale = 1.2)

# Option 2, with food group plot on its own row aligned to the full plot area
library(cowplot)
ggdraw() +
  draw_plot(a, x = 0, y = .5, width = .5, height = .5) +
  draw_plot(b, x = .5, y = .5, width = .5, height = .5) +
  draw_plot(c_with_icons + theme(legend.position = "none"), x = 0, y = 0, width = 1, height = 0.5) +
  draw_plot_label(label = c("a", "b", "c"), size = 15,
                  x = c(0, 0.5, 0), y = c(1, 1, 0.5))
ggsave(here("output", "figs", "declines_ms", "fig_correlation_results_v2.png"), height = 6.5, width = 8, dpi = 600, scale = 1.2)
```

## SI figure showing raw correlation coefficients

```{r}
cor_results %>% 
  filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
  select(-total_supply_per_cap, 
         -stock_var_per_cap,
         -feed_per_cap,
         -seed_per_cap) %>% 
  pivot_longer(cols = where(is.double), names_to = "element", values_to = "rho") %>% 
  
  ggplot(aes(x = rho, fill = factor(element, levels = c("prod_per_cap", "import_per_cap", "export_per_cap", "feed_per_cap", "seed_per_cap")), color = factor(element, levels = c("prod_per_cap", "import_per_cap", "export_per_cap", "feed_per_cap", "seed_per_cap")))) +
  geom_density(size = 1.25, alpha = 0.1) +
  scale_color_manual(values = c(paletteer_d(pal)[2:6]), 
                     labels = c("export_per_cap" = "Exports", 
                                "import_per_cap" = "Imports",
                                "prod_per_cap" = "Production",
                                "feed_per_cap" = "Feed",
                                "seed_per_cap" = "Seed")) +
  scale_fill_manual(values = c(paletteer_d(pal)[2:6]), 
                    labels = c("export_per_cap" = "Exports", 
                               "import_per_cap" = "Imports",
                               "prod_per_cap" = "Production",
                               "feed_per_cap" = "Feed",
                               "seed_per_cap" = "Seed")) +
  scale_x_continuous(breaks = seq(-1, 1, by = 0.25), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0))) +
  labs(x = "Spearman's rho",
       y = "Density",
       color = bquote(Supply~element~(kg~capita^-1~yr^-1)),
       fill = bquote(Supply~element~(kg~capita^-1~yr^-1))) +
  cowplot::theme_half_open() +
  theme(legend.position = c(0.03, 0.85),
        plot.margin = unit(c(7,17,7,7), "pt"))

ggsave(here("output", "figs", "declines_ms", "fig_correlation_results_si.png"), height = 6, width = 10, dpi = 600)
```

## SI plot showing sensitivity of rho

How sensitive are the dominant driver results to the choice of rho?

```{r}
main_correlates_sensitivity <- cor_results %>% 
  filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Unrecovered declines only
  # drop_na(ends_with("per_cap")) %>% # Remove declines that have no rho value for ANY correlate (this is due to missing data) -- decided against this, very few have complete data
  select(-total_supply_per_cap,
         -feed_per_cap,
         -seed_per_cap, # Seed positively correlated with food supply
         -stock_var_per_cap) %>% # Stock variation very weakly correlated to supply 
  pivot_longer(cols = ends_with("per_cap"), names_to = "element", values_to = "rho") %>% 
  mutate(rho = abs(rho)) %>% # Take absolute value of rho
  group_by(decline_id) %>% 
  summarize(dominant_correlate_0.9 = if_else(max(na.omit(rho)) >= 0.9, 
                                         paste(element[which(rho >= 0.9)], collapse = ", "),
                                         paste(element[which(rho == max(na.omit(rho)))], collapse = ", ")),
            dominant_correlate_0.8 = if_else(max(na.omit(rho)) >= 0.8, 
                                         paste(element[which(rho >= 0.8)], collapse = ", "),
                                         paste(element[which(rho == max(na.omit(rho)))], collapse = ", ")),
            dominant_correlate_0.7 = if_else(max(na.omit(rho)) >= 0.7, 
                                         paste(element[which(rho >= 0.7)], collapse = ", "),
                                         paste(element[which(rho == max(na.omit(rho)))], collapse = ", ")),
            dominant_correlate_0.6 = if_else(max(na.omit(rho)) >= 0.6, 
                                         paste(element[which(rho >= 0.6)], collapse = ", "),
                                         paste(element[which(rho == max(na.omit(rho)))], collapse = ", ")),
            dominant_correlate_0.5 = if_else(max(na.omit(rho)) >= 0.5, 
                                         paste(element[which(rho >= 0.5)], collapse = ", "),
                                         paste(element[which(rho == max(na.omit(rho)))], collapse = ", "))) %>%
  mutate(across(starts_with("dominant_correlate_"), ~if_else(.x == "", NA_character_, .x))) %>% 
  ungroup() %>% 
  left_join(select(declines, decline_id, area, area_code, item, item_code, food_group, sector), by = "decline_id") # Rejoin food data

main_correlates_sensitivity %>% 
  na.omit() %>% 
  pivot_longer(cols = starts_with("dominant_correlate_"), names_to = "rho_thresh", values_to = "dominant_correlate") %>% 
  mutate(rho_thresh = parse_number(rho_thresh)) %>% 
  separate_rows(dominant_correlate, sep = ", ") %>% 
  count(rho_thresh, dominant_correlate) %>% 
  mutate(prop = n/nrow(na.omit(main_correlates_sensitivity))) %>% 
  
  ggplot(aes(x = rho_thresh, fill = factor(dominant_correlate, levels = c("prod_per_cap", "import_per_cap", "export_per_cap")))) +
  geom_col(aes(x = rho_thresh, y = prop), position = "dodge") + 
  labs(y = "Percentage of supply declines",
       x = "Spearman's rho threshold",
       fill = "Dominant correlate") +
  scale_fill_manual(values = c(paletteer_d(pal)[2:5]),
                    labels = c("export_per_cap" = "Exports",
                               "import_per_cap" = "Imports",
                               "prod_per_cap" = "Production")
  ) +
  scale_x_continuous(breaks = c(0.5, 0.6, 0.7, 0.8, 0.9)) +
  scale_y_continuous(limits = c(0, 0.8), 
                     breaks = seq(0, 0.8, by = 0.1),
                     labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0))) +
  cowplot::theme_half_open() +
  theme(plot.margin = unit(c(7,17,7,7), "pt"),
        legend.position = "top")

ggsave(here("output", "figs", "declines_ms", "fig_correlation_sensitivity_si.png"), height = 6, width = 8, dpi = 600)
```


## Exploring set plots
```{r}
# # Set plots (Test out a few packages for making venn diagrams and upset plots)
# venn_input <- main_correlates %>% 
#   count(dominant_correlate) %>% 
#   # mutate(dominant_correlate = case_when(dominant_correlate == "export_per_cap" ~ "Exports", 
#   #                                       dominant_correlate == "import_per_cap" ~ "Imports", 
#   #                                       dominant_correlate == "prod_per_cap" ~ "Production",
#   #                                       dominant_correlate == "prod_per_cap, export_per_cap" ~ "Production, Exports",
#   #                                       dominant_correlate == "prod_per_cap, import_per_cap" ~ "Production, Imports",
#   #                                       dominant_correlate == "import_per_cap, export_per_cap" ~ "Imports, Exports",
#   #                                       dominant_correlate == "prod_per_cap, import_per_cap, export_per_cap" ~ "Production, Imports, Exports",
#   #                                       TRUE ~ dominant_correlate)) %>% 
#   mutate(dominant_correlate = str_replace_all(dominant_correlate, ", ", "&")) %>% 
#   deframe()
# 
# library(UpSetR)
# (upset <- upset(fromExpression(venn_input), 
#       order.by = "freq",
#       mainbar.y.label = "Set size (# declines)", 
#       sets.x.label = "Intersection size (# declines)",
#       text.scale = 2)
# )
# 
# png(here::here("output", "figs", "declines_ms", "fig_correlation_results_upset.png"), height = 5, width = 7, unit = "in", res = 300)
# upset
# dev.off()
# 
# library(eulerr)
# (venn <- plot(euler(venn_input), 
#      fills = paletteer_d(pal),
#      alpha = 0.75,
#      edges = FALSE,
#      quantities = TRUE) # Missing some small sets
# )
# 
# euler(venn_input) # A large amount of sets --> very imperfect solution
# euler(venn_input, shape = "ellipse") # Fitting to ellipse doesn't help
# 
# 
# png(here::here("output", "figs", "declines_ms", "fig_correlation_results_venn.png"), height = 5, width = 5, unit = "in", res = 300)
# venn
# dev.off()
```

# Investigate production-associated declines

How many declines that were associated with per capita reduction in production experienced actual reductions in gross production?

First, extract the production-associated declines:
```{r}
# Checking count
# main_correlates %>% 
#   filter(decline_id %in% declines$decline_id[declines$recovery_0.95 == FALSE]) %>% # Select unrecovered declines 
#   separate_rows(dominant_correlate, sep = ", ") %>% 
#   filter(dominant_correlate == "prod_per_cap") %>% 
#   count(decline_id, sort = TRUE)
  
declines_assoc_w_prod <- main_correlates %>% 
  separate_rows(dominant_correlate, sep = ", ") %>% 
  filter(dominant_correlate == "prod_per_cap") # n = 842
```

Get production time series during those declines.
```{r}
food_production <- food_balance %>% 
  
  # Select production only
  filter(element_code == 5511) %>% 
  mutate(production = quantity*1000*1000, .keep = "unused") %>% # Convert to kg
  select(-unit, -element, -element_code) %>% 
  
  # Filter to only the years within unrecovered decline events associated with per cap production
  right_join(
    ts_w_declines %>% 
      select(decline_id, area_code, area, item_code, item, year) %>% 
      filter(decline_id %in% declines_assoc_w_prod$decline_id),
    by = c("area_code", "area", "item_code", "item", "year")
    ) %>%

  arrange(decline_id, year) 
```

For every decline, extract coefficient of linear model: `production ~ year`
```{r}
# Overall
food_production %>% 
  nest(decline_ts = -decline_id) %>% 
  mutate(lm = map(decline_ts, ~lm(production ~ year, data = .))) %>% 
  mutate(lm_tidied = map(lm, tidy)) %>% 
  unnest(lm_tidied) %>% 
  filter(term == "year") %>% 
  select(decline_id, 
         lm_coef = estimate) %>% # No NAs generated at this stage..
  right_join(declines_assoc_w_prod) %>%
  count(abs_prod_decline = lm_coef < 0) %>% 
  mutate(prop = n/sum(n))

# By food group
food_production %>% 
  nest(decline_ts = -decline_id) %>% 
  mutate(lm = map(decline_ts, ~lm(production ~ year, data = .))) %>% 
  mutate(lm_tidied = map(lm, tidy)) %>% 
  unnest(lm_tidied) %>% 
  filter(term == "year") %>% 
  select(decline_id, 
         lm_coef = estimate) %>% # No NAs generated at this stage..
  right_join(declines_assoc_w_prod) %>%
  count(abs_prod_decline = lm_coef < 0, 
        food_group) %>% 
  mutate(prop = n/sum(n))

food_production %>% 
  nest(decline_ts = -decline_id) %>% 
  mutate(lm = map(decline_ts, ~lm(production ~ year, data = .))) %>% 
  mutate(lm_tidied = map(lm, tidy)) %>% 
  unnest(lm_tidied) %>% 
  filter(term == "year") %>% 
  select(decline_id, 
         lm_coef = estimate) %>% 
  right_join(declines_assoc_w_prod) %>%
  count(abs_prod_decline = lm_coef < 0, 
        food_group) %>% 
  
  # Plot
  ggplot(aes(x = n, y = food_group, fill = abs_prod_decline)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(fill = "Absolute production decline?") +
  scale_x_continuous(labels = scales::percent) +
  cowplot::theme_cowplot()
```

## Check if available data meets a threshold of supply

```{r}
ts_w_declines %>% 
  filter(decline_id %in% unique(declines_assoc_w_prod$decline_id)) %>%
  rename(supply_per_cap = quantity) %>%
  left_join(food_balance_per_cap) %>% 
  select(decline_id, year, supply_per_cap, prod_per_cap, export_per_cap, import_per_cap, feed_per_cap, seed_per_cap) %>% 
  group_by(decline_id) %>% 
  filter(year == max(year)) %>% 
  ungroup() %>% 
  mutate(export_per_cap = export_per_cap * -1,
         feed_per_cap = feed_per_cap * -1,
         seed_per_cap = seed_per_cap * -1) %>% 
  rowwise() %>% 
  mutate(sum = sum(c_across(4:8), na.rm = TRUE)) %>% 
  mutate(prop_of_supply = sum/supply_per_cap) %>% 
  mutate(within_thresh = if_else(prop_of_supply < 1.15 & prop_of_supply > .85, TRUE, FALSE)) %>% 
  count(within_thresh)
```

## Use same method to investigate if harvestable area and yield also declined

Get area & yield time series during production-associated declines of CROP items only. Note that items in the crops dataset are not the same as those in food balance sheets, so we'll need to map across them. Mel provided a key she made that does so. In it, FAO_item_code represents the crop item code, and FAO_item_code_NFB represents the FBS item code.
```{r}
distinct(crops, element_code, element)

# Area harvested during declines
area_harvested <- crops %>% 
  
  # Select production only
  filter(element_code == 5312) %>% # hectares
  select(-unit, -element, -element_code, -flag) %>% 
  
  # Add FAO item from key
  left_join(fbs_to_crops_key %>% 
              select(FAO_item_code, FAO_item_code_NFB),
            by = c("item_code" = "FAO_item_code")) %>% 
    
  # Summarize all crop items by FBS item
  group_by(area_code, area, FAO_item_code_NFB, year) %>% 
  summarize(value = sum(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  
  # Filter to only the years within unrecovered CROP decline events associated with per cap production
  right_join(
    ts_w_declines %>% 
      select(decline_id, area_code, area, item_code, item, sector, year) %>% 
      filter(decline_id %in% declines_assoc_w_prod$decline_id,
             sector == "Crops"),
    by = c("area_code", "area", "FAO_item_code_NFB" = "item_code", "year")
    ) %>%

  arrange(decline_id, year) 

# Yield during declines
yield <- crops %>% 
  
  # Select production only
  filter(element_code == 5419) %>% # hectograms/hectare
  select(-unit, -element, -element_code, -flag) %>% 
  
  # Add FAO item from key
  left_join(fbs_to_crops_key %>% 
              select(FAO_item_code, FAO_item_code_NFB),
            by = c("item_code" = "FAO_item_code")) %>% 
    
  # Summarize all crop items by FBS item
  group_by(area_code, area, FAO_item_code_NFB, year) %>% 
  summarize(value = sum(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  
  # Filter to only the years within unrecovered CROP decline events associated with per cap production
  right_join(
    ts_w_declines %>% 
      select(decline_id, area_code, area, item_code, item, sector, year) %>% 
      filter(decline_id %in% declines_assoc_w_prod$decline_id,
             sector == "Crops"),
    by = c("area_code", "area", "FAO_item_code_NFB" = "item_code", "year")
    ) %>%

  arrange(decline_id, year) 
```

For every decline, extract coefficient of linear model: `value ~ year`, where value == area or yield
```{r}
# Overall
area_harvested_results <- area_harvested %>% 
  na.omit() %>% 
  nest(decline_ts = -decline_id) %>% 
  mutate(lm = map(decline_ts, ~lm(value ~ year, data = .))) %>% 
  mutate(lm_tidied = map(lm, tidy)) %>% 
  unnest(lm_tidied) %>% 
  filter(term == "year") %>% 
  select(decline_id, 
         lm_coef = estimate) %>% # No NAs generated at this stage..
  left_join(declines_assoc_w_prod) 

area_harvested_results %>%
  count(harvestable_area_decline = lm_coef < 0) %>% 
  mutate(prop = n/sum(n))

# By food group
yield_results <- yield %>% 
  na.omit() %>% 
  nest(decline_ts = -decline_id) %>% 
  mutate(lm = map(decline_ts, ~lm(value ~ year, data = .))) %>% 
  mutate(lm_tidied = map(lm, tidy)) %>% 
  unnest(lm_tidied) %>% 
  filter(term == "year") %>% 
  select(decline_id, 
         lm_coef = estimate) %>% # No NAs generated at this stage..
  left_join(declines_assoc_w_prod)

yield_results %>% 
count(yield_decline = lm_coef < 0) %>% 
  mutate(prop = n/sum(n))

# What percentage of production-associated crop declines also had declines in harvestable area while yield stayed the same or increased?
yield_results %>% 
  select(decline_id, yield_coef = lm_coef) %>% 
  left_join(area_harvested_results %>% 
              select(decline_id, area_coef = lm_coef)) %>% 
  filter(area_coef < 0,
         yield_coef >= 0) %>% 
  nrow() / 519 
```
